Concurrent Programming with Harmony

### 2020-07-15 book/paper.tex
Clarification
Typo

A method \texttt{m} with argument $a$ is invoked in its
most basic form as follows (assigning the result to~\texttt{r}).
\begin{code}
$r$ = m $a$;
\end{code}
That's right, no parentheses are required.  In fact, if you invoke
\texttt{m}($a$), the argument is ($a$), which is the same
as $a$.
If you invoke \texttt{m()}, the argument is \texttt{()},
which is the empty tuple.
If you invoke \texttt{m}($a$, $b$), the argument is ($a$, $b$),
the tuple consisting of values $a$ and $a$.

@@

\section*{Exercises}
\begin{problems}
\item Implement an atomic swap operation.  It should take two pointer arguments
and swap the values.
\item \label{ex:swap} Implement a spinlock using the atomic swap operation.
\item For the solution to \autoref{ex:swap},
write out the invariants that need to hold and check them using Harmony.
\item People who use an ATM often first check their balance and then withdraw
a certain amount of money not exceeding their balance.  A negative balance
is not allowed.  \autoref{fig:atm} shows two operations on bank accounts:
one to check the balance and one to withdraw money.
Note that all operations on accounts are carefully protected by a lock
(i.e., there are no data races).
The \texttt{customer}
method models going to a particular ATM and withdrawing money not exceeding
the balance.  Method \texttt{checker} looks for negative balances.
Running the code through Harmony reveals that there is a bug.
It is a common type of concurrency bug known as \emph{Time Of Check Time Of
Execution} (TOCTOE).
\index{Time Of Check Time Of Execution}
\index{TOCTOE}
In this case, by the time the withdraw operation is performed,
the balance can have changed.
\begin{enumerate}
\item Fix the code in \autoref{fig:atm}.
\item Is it necessary to obtain a lock in \texttt{atm\_check\_balance()}?
\end{enumerate}
\end{problems}

>>>>

A method \texttt{m} with argument $a$ is invoked in its
most basic form as follows (assigning the result to~\texttt{r}).
\begin{code}
$r$ = m $a$;
\end{code}
That's right, no parentheses are required.  In fact, if you invoke
\texttt{m}($a$), the argument is ($a$), which is the same
as $a$.
If you invoke \texttt{m()}, the argument is \texttt{()},
which is the empty tuple.
If you invoke \texttt{m}($a$, $b$), the argument is ($a$, $b$),
the tuple consisting of values $a$ and $b$.

@@

\section*{Exercises}
\begin{problems}
\item Implement an atomic swap operation.  It should take two pointer arguments
and swap the values.
\item \label{ex:swap} Implement a spinlock using the atomic swap operation.
\item For the solution to \autoref{ex:swap},
write out the invariants that need to hold and check them using Harmony.
\item People who use an ATM often first check their balance and then withdraw
a certain amount of money not exceeding their balance.  A negative balance
is not allowed.  \autoref{fig:atm} shows two operations on bank accounts:
one to check the balance and one to withdraw money.
Note that all operations on accounts are carefully protected by a lock
(i.e., there are no data races).
The \texttt{customer}
method models going to a particular ATM and withdrawing money not exceeding
the balance.  Method \texttt{checker} looks for negative balances.
Running the code through Harmony reveals that there is a bug.
It is a common type of concurrency bug known as \emph{Time Of Check Time Of
Execution} (TOCTOE).
\index{Time Of Check Time Of Execution}
\index{TOCTOE}
In this case, by the time the withdraw operation is performed,
the balance can have changed.
\begin{enumerate}
\item Fix the code in \autoref{fig:atm}.  Note, you should leave the
customer code the same. You are only allowed to change the
\texttt{atm\_} methods, and you cannot use the \textbf{atomic}
statement.
\item Is it necessary to obtain a lock in \texttt{atm\_check\_balance()}?
\end{enumerate}
\end{problems}
